Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-07-20 analysis/paper/sim_survey_paper.Rmd
Start results section (add first paragraph)

Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project. Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g. bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997]. As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland. Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol. Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par). A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests). Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$. We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:

$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})}{N_a N_y N_k}}$$
Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e. catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$). All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).

@@

- add population trend plot?

\@ref(fig:sp-fig)

@@

In addition to the simulation run presented in body of this paper, which simulates surveys over a population that forms age-specific clusters, we also simulated surveys over a population that assumes the same spatial clustering across all age groups. There is therefore no intra-haul correlation in this series of surveys. Aside from this change to the spatial distribution settings, all other settings were kept the same. Results are presented below in Figures A1, A2 and A3. Aside from systematic underestimation in the low set density scenarios, there are little signs of clear bias in the stratified estimates. The bias present under the low set density scenario is likely related to the disproportionate sampling of the narrow shallow and deep strata in the survey grid. The lack of bias from the other scenarios help verify the issues introduced by age-specific clustering to stratified estimates of abundance at age. These results also indicate that little precision is gained by increasing length or age sampling effort beyond 100 lengths per set and 10 ages per length group, respectively. Given this result, it is probably not unreasonable to speculate that such sampling effort helps define the shape of the age distribution, but extra effort does little to improve estimates of abundance at age; such gains are only obtained by increasing set density.

>>>>

Ideally the parameter settings used here would be based on model fits to real data, however, conditioning the operating model described above was outside the scope of this project. Moreover, many of the parameters described above may not be identifiable or it may be difficult to obtain un-biased estimates [e.g. bias introduced by length-stratified sampling complicate the fitting of growth models; @Morgan1997]. As such, parameters were informally selected to roughly correspond to observations typically obtained by the Fisheries and Oceans Canada survey of cod in NAFO division 3Ps, Newfoundland. Though this approach is far from exact, the goal was to approximate the life-history and distribution of an actual population to simulate an arena for testing the consequences of surveys with different sampling protocol. Parameter settings and descriptions used in this case study are presented in Table \@ref(tab:par). A series of `r nrow(sim$surveys)` surveys were tested using an array of different sampling protocol outlined in Table \@ref(tab:tests). Each survey was simulated `r max(sim$age_strat_error$sim)` times over the same population and samples were run through a stratified analysis [@Smith1981] to obtain an estimate of abundance at age for each year for each simulation, $\hat{I}_{a,y,k}$. Intra-haul correlation (i.e. intraclass correlation) of the set catch was calculated by fitting a mixed-model to the lengths data with set as a random effect [fit using the lme4 package; @Bates2015]; variance estimates from this model were then plugged into the ICC(1) formula presented in McGraw and Wong [-@Mcgraw1996]. We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance estimates from each survey:

$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})}{N_a N_y N_k}}$$
Where $N_a$, $N_y$, and $N_k$ are the number of years, ages and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e. catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$). All simulations and analyses were conducted using `r substr(version$version.string, 1, 15)` [@R2017] and the code has been developed into a package called `SimSurvey` to facilitate future work and collaboration (https://github.com/PaulRegular/SimSurvey).

@@

```{r icc, cache = TRUE}
len_samp <- merge(sim$setdet, sim$samp, by = "set")
len_samp <- len_samp[len_samp$sim == 1 & len_samp$measured, ]
sim_icc <- round(icc(len_samp$length, len_samp$set), 1)
sim_icc <- sprintf("%.2f", sim_icc)
real_icc <- "0.70" # calculated in the file "imatate_cod_data.R"
```

A dynamic population that varies across ages, time and space is simulated with the parameter settings used here (Table \@ref(tab:par)). Figure \@ref(fig:sp-fig) depicts the aggregate distribution of the population in one year of the simulation and the corresponding catches of one stratified random survey over this population. Layered within this, of course, is the distributions of each age, as well as age and length sampling details. Though age-specific details are not depicted for simplicity, the general pattern of spatial aggregation apparent in Figure \@ref(fig:sp-fig) is similar across all ages. That is, all ages tend to aggregate in fairly dense clusters and low-density zones are fairly wide ranging. The survey, therefore, tends to be characterized by regions with large catches interspersed with regions of no catch (Figure \@ref(fig:sp-fig)). The samples within each successful set are also dynamic and rarely include all age groups. This is because only moderate correlation was imposed on the distributions of ages 1-5+, allowing individuals of different ages to occupy different locations. This results in an intra-haul correlation of approximately `r sim_icc`, which, in this case, is of equal magnitude to the intra-haul correlation observed across more than 20 years of survey data of 3Ps cod.

Next discuss trend...

@@

In addition to the run presented in body of this paper, which simulates surveys over a population that forms age-specific clusters, we also simulated surveys over a population that assumes the same spatial clustering across all age groups. There is therefore no intra-haul correlation in this series of surveys. Aside from this change to the spatial distribution settings, all other settings were kept the same. Results are presented below in Figures A1, A2 and A3. Aside from systematic underestimation in the low set density scenarios, there are little signs of clear bias in the stratified estimates. The bias present under the low set density scenario is likely related to the disproportionate sampling of the narrow shallow and deep strata in the survey grid. The lack of bias from the other scenarios help verify the issues introduced by age-specific clustering to stratified estimates of abundance at age. These results also indicate that little precision is gained by increasing length or age sampling effort beyond 100 lengths per set and 10 ages per length group, respectively. Given this result, it is probably not unreasonable to speculate that such sampling effort helps define the shape of the age distribution, but extra effort does little to improve estimates of abundance at age; such gains are only obtained by increasing set density.
